\documentclass[12pt, a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage{geometry}
\geometry{a4paper, margin=1in}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{titlesec}
\usepackage{xcolor}

\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,
    urlcolor=cyan,
    pdftitle={Big Data Transit System Documentation Outline}
}

\title{\textbf{System Documentation Outline}\\
\large Big Data Transit System}
\date{\vspace{-5ex}}

\begin{document}

\maketitle
\tableofcontents
\newpage

\section{Executive Summary}
\begin{itemize}[leftmargin=*]
    \item One-paragraph overview of the system, problem solved, and scale.
    \item Key outcomes: real-time tracking, analytics, and reliability goals.
\end{itemize}

\section{System Goals and Constraints}
\begin{itemize}[leftmargin=*]
    \item Scale targets (buses, messages per second, daily volume).
    \item Latency and availability targets (edge to UI).
    \item Offline-first requirements and retry behavior.
    \item Data retention and storage growth assumptions.
\end{itemize}

\section{Architecture Overview}
\subsection{Component Map}
\begin{itemize}[leftmargin=*]
    \item Edge simulator (Bus module).
    \item Core ingestion server.
    \item Website backend (BFF).
    \item Website frontend.
    \item Simulation orchestrator.
\end{itemize}

\subsection{Deployment Topology}
\begin{itemize}[leftmargin=*]
    \item Environments (local, staging, production).
    \item Ports, protocols, and network boundaries.
    \item External services and dependencies.
\end{itemize}

\section{Data Contracts and Schemas}
\subsection{Edge Telemetry Payload}
\begin{itemize}[leftmargin=*]
    \item GPS payload fields and types.
    \item Optional fields (edge anomalies, quality flags).
    \item Example JSON payload.
\end{itemize}

\subsection{Server API Contracts}
\begin{itemize}[leftmargin=*]
    \item Request and response schemas for write and read APIs.
    \item Error responses and status codes.
\end{itemize}

\subsection{Database Schema Map}
\begin{itemize}[leftmargin=*]
    \item Raw telemetry tables.
    \item Derived analytics tables.
    \item Metrics tables.
\end{itemize}

\section{Edge Simulator (Bus Module)}
\subsection{Runtime Architecture}
\begin{itemize}[leftmargin=*]
    \item Bus lifecycle, GPS loop, Sender, Receiver, Database.
    \item Threading model and concurrency limits.
\end{itemize}

\subsection{Data Generation}
\begin{itemize}[leftmargin=*]
    \item Route loading and waypoint iteration.
    \item Snapshot cadence and timestamp strategy.
\end{itemize}

\subsection{Local Persistence and Offline Behavior}
\begin{itemize}[leftmargin=*]
    \item SQLite schema, WAL mode, and sync flags.
    \item Retry strategy and backlog handling.
\end{itemize}

\subsection{Network Transmission}
\begin{itemize}[leftmargin=*]
    \item Persistent socket behavior.
    \item HTTP request format and keep-alive handling.
\end{itemize}

\subsection{Command Channel}
\begin{itemize}[leftmargin=*]
    \item Receiver protocol, command types, and handling.
\end{itemize}

\section{Core Ingestion Server}
\subsection{Startup Sequence}
\begin{itemize}[leftmargin=*]
    \item Environment loading, schema init, cache warm-up, metrics flusher.
    \item Failure modes and startup dependencies.
\end{itemize}

\subsection{Write API: GPS Ingestion}
\subsubsection{Endpoint and Contract}
\begin{itemize}[leftmargin=*]
    \item Route, method, required fields.
    \item Example request payload.
\end{itemize}

\subsubsection{Validation Rules}
\begin{itemize}[leftmargin=*]
    \item Required field checks.
    \item Type validation and bounds checks.
    \item Duplicate detection strategy.
    \item Edge anomaly flags and skip rules.
\end{itemize}

\subsubsection{Acknowledgement Strategy}
\begin{itemize}[leftmargin=*]
    \item Immediate response behavior.
    \item Async processing and error logging.
\end{itemize}

\subsection{Processing Pipeline}
\begin{itemize}[leftmargin=*]
    \item Vehicle state tracking.
    \item Raw telemetry insert.
    \item Latest state upsert.
    \item Anomaly detection.
    \item Stop and dwell event generation.
    \item Headway and speed profile generation.
    \item Trip lifecycle updates.
\end{itemize}

\subsection{Read APIs (Analytics and Live)}
\begin{itemize}[leftmargin=*]
    \item Endpoint list and routes.
    \item Query parameters and validation rules.
    \item Response formats and pagination.
\end{itemize}

\section{Data Validation and Quality}
\begin{itemize}[leftmargin=*]
    \item Edge validation checks before send.
    \item Server-side validation and error handling.
    \item Out-of-order and duplicate packet handling.
    \item Quality flags and anomaly tagging.
\end{itemize}

\section{Database Design}
\subsection{Raw Telemetry}
\begin{itemize}[leftmargin=*]
    \item Table purpose and key columns.
    \item Uniqueness constraints and indexes.
\end{itemize}

\subsection{Derived Analytics Tables}
\begin{itemize}[leftmargin=*]
    \item stop\_events, dwell\_times, headway\_records.
    \item speed\_profiles, trip\_logs, anomaly\_events.
\end{itemize}

\subsection{Ingestion Metrics}
\begin{itemize}[leftmargin=*]
    \item ingest\_metrics\_minute schema.
    \item Phase naming and usage.
\end{itemize}

\section{Website Backend (BFF)}
\subsection{Responsibilities}
\begin{itemize}[leftmargin=*]
    \item Authentication and session management.
    \item Rate limiting and request shaping.
    \item Caching strategy and data aggregation.
\end{itemize}

\subsection{API Surface}
\begin{itemize}[leftmargin=*]
    \item Endpoints exposed to frontend.
    \item Request validation and error handling.
\end{itemize}

\section{Website Frontend}
\subsection{Data Flow}
\begin{itemize}[leftmargin=*]
    \item Data fetching strategy (polling, streaming, or hybrid).
    \item State management and update cadence.
\end{itemize}

\subsection{Visualization}
\begin{itemize}[leftmargin=*]
    \item Map base layer (MapLibre).
    \item GPU layers (deck.gl) for vehicles, routes, and stations.
    \item Event overlays (anomalies, dwell, headway).
\end{itemize}

\subsection{Performance and UX}
\begin{itemize}[leftmargin=*]
    \item Throttling and batching.
    \item Viewport culling and level-of-detail.
    \item Interaction and animation rules.
\end{itemize}

\section{End-to-End Data Flow}
\begin{enumerate}[leftmargin=*]
    \item GPS snapshot created on bus.
    \item HTTP POST to ingestion server.
    \item Validation and async processing.
    \item Database writes and derived events.
    \item Read API query.
    \item Frontend visualization update.
\end{enumerate}

\section{Observability and Monitoring}
\begin{itemize}[leftmargin=*]
    \item Logs and error reporting.
    \item Metrics for ingest performance and drop rate.
    \item Dashboards and alert thresholds.
\end{itemize}

\section{Deployment and Configuration}
\begin{itemize}[leftmargin=*]
    \item Environment variables and secrets.
    \item Config files and runtime flags.
    \item Service startup order.
\end{itemize}

\section{Testing Strategy}
\begin{itemize}[leftmargin=*]
    \item Unit tests for validation and processing rules.
    \item Integration tests for ingest and read APIs.
    \item Load tests and benchmarks.
\end{itemize}

\section{Security and Privacy}
\begin{itemize}[leftmargin=*]
    \item Authentication and authorization.
    \item Rate limiting and abuse prevention.
    \item Data minimization and retention.
\end{itemize}

\section{Limitations and Roadmap}
\begin{itemize}[leftmargin=*]
    \item Current bottlenecks and known gaps.
    \item Kafka, Redis, and Spark adoption plan.
    \item Edge enhancements and replay pipeline.
\end{itemize}

\end{document}
